Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-07-18 main.tex
minor tweak
Update on Overleaf.
rearrange

% DONE: MV : Be more didactical: write following as the existence of a potential. 
Given a domain $\Omega \subseteq \bbR^{n}$ and Lebesgue exponent $1 \leq p \leq \infty$,
we search for a constant $C_{\grad,\Omega,p} > 0$ such that the following holds:
for every gradient vector field $\bff \in \nabla W^{1,p}(\Omega)$ there exists a scalar potential $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:intro:inequalitypf:gradient}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align}
We call this inequality \emph{Poincar\'e--Friedrichs inequality} and the constant $C_{\grad,\Omega,p}$ is called the \emph{Poincar\'e--Friedrichs constant} with exponent $p$. 
The different scalar potentials only differ by a constant. 
Therefore, %when the domain is connected, then 
this inequality is equivalent to the following statement: 
\begin{align}\label{math:intro:poincarefriedrichs:grad}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    \forall
    u \in W^{1,p}(\Omega) % TODO: Use \forall notation 
    .\tag{PF} 
\end{align}
% Given a domain $\Omega \subseteq \bbR^{n}$ and Lebesgue exponent $1 \leq p \leq \infty$,
% we search for a constant $C_{\grad,\Omega,p} > 0$ such that 
% \begin{align}\label{math:intro:poincarefriedrichs:grad}
%     \min\limits_{c\in\bbR}
%     \| u - c \|_{L^{p}(\Omega)}
%     \leq 
%     C_{\grad,\Omega,p} \| \nabla u \|_{L^p(\Omega)},
%     \quad 
%     u \in W^{1,p}(\Omega) % TODO: Use \forall notation 
%     .\tag{PF} 
% \end{align}
Finding a potential of a given gradient $\nabla u$ is generally a nonlinear operation. 
The operator norm of any \emph{linear} potential construction is an upper bound for the Poincar\'e--Friedrichs constant. 
% Even though the existence of such a bounded linear potential construction is not evident a priori, picking the average-free potential 
One example for a bounded linear operator is the average-free potential:
\begin{align*}
    \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the constant $C_{{\rm P},\Omega,p} > 0$ in the Poincar\'e inequality 
\begin{align}\label{math:intro:poincare:grad}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm P},\Omega,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    , \tag{P} % DONE: tag poincare inequality 
\end{align}
where $u_\Omega$ denotes the average of $u$.
Considerable research effort has gone into finding (or estimating) the optimal constants in the Poincar\'e--Friedrichs inequalities~\eqref{math:intro:poincarefriedrichs:grad} or Poincar\'e inequalities~\eqref{math:intro:poincare:grad}.

@@

% DONE: INTRODuce a reference simplex 
% DONE: DEFINE THE JACOBIAN 
\todo{MWL: If we need to define the Jacobian formally, where should that be put? Maybe an extra paragraph \textbf{Notation} immediately at the end of the introduction section.}

@@

\begin{lemma}    
    Suppose that an $n$-dimensional manifold triangulation $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$.
    For $0 \leq m < M$, write 
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m},
        \qquad 
        I_{m} := X_{m} \cap T_{m+1}.
    $$ 
    Then $I$ is the union of $k$ different facets of $T$, 
    and their intersection is a simplex $S \in \calT$ of dimension $n-k$.
    Moreover, 
    $$
        \patch_{X_{m+1}}(S) = \patch_{\calT}(S).
    $$
\end{lemma}
\begin{proof}
    We know $I_{m}$ is a submanifold of the boundary of $T$, and so it must be a collection of $k$ facets
    that constitute the local patch around some simplex $S$ of dimension $n-k$.
\end{proof}

\begin{theorem}
    Let $\calT$ be an $n$-dimensional triangulation and let $S \in \calT$ be a simplex of dimension $k < n$.
    Let $T \in \patch_{\calT}(S)$ be of dimension $n$ and let $M := |\patch_{\calT}(S)| \setminus T$. 
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \varphi : T \rightarrow M
    \end{align*}
    satisfying
    \begin{align*}
        \Lip( \varphi,      T ) \leq \Cvier{n}{k},
        \quad 
        \Lip( \varphi^{-1}, M ) \leq \Cfive{n}{k},
    \end{align*}
    where $\Cvier{n}{k} \leq ...$ and $\Cfive{n}{k} \leq ...$
\end{theorem}

@@

\subsection{Inequalities for partial boundary conditions on a single simplex}

\subsection{Constructions over shellable triangulations}

\subsection{Numerical examples}

@@

% https://mathoverflow.net/questions/232302/lipschitz-constant-of-cental-projection-of-unit-ball-to-surface-of-convex-body
% https://eudml.org/doc/258483
% https://math.stackexchange.com/questions/1476085/uniformly-convex-approximation-of-convex-domain
% SMOOTH EXHAUSTION FUNCTIONS IN CONVEX DOMAINS
\begin{lemma}
    Let $K \subseteq \bbR^{n}$ be a compact set star-shaped with respect to the origin 
    that contains the ball $B_{r}(0)$ and that is contained in the ball $B_{R}(0)$. Then the function 
    \begin{align*}
        f : S \rightarrow \partial K, \quad x \mapsto \arg( \partial K \cap \bbR x )
    \end{align*}
    satisfies 
    \begin{align*}
        R^{2}/r % euclidean 
    \end{align*}
    If $K$ is convex, then 
    \begin{align*}
        g : S \rightarrow \bbR, \quad x \mapsto \| \bbR x \cap \partial K \|
    \end{align*}
    satisfies 
    \begin{align*}
        R \sqrt{ R^{2} / r^{2} - 1 } % angular distance 
        \\
        \frac R r \sqrt{ R^{2} - r^{2} } % angular distance 
    \end{align*}
\end{lemma}

\begin{lemma}
    .
\end{lemma}

>>>>

% DONE: MV : Be more didactical: write following as the existence of a potential. 
Given a domain $\Omega \subseteq \bbR^{n}$ and Lebesgue exponent $1 \leq p \leq \infty$,
we search for a constant $C_{\grad,\Omega,p} > 0$ such that the following holds:
for every gradient vector field $\bff \in \nabla W^{1,p}(\Omega)$ there exists a scalar potential $u \in W^{1,p}(\Omega)$
such that $\nabla u = \bff$ and 
\begin{align}\label{math:intro:inequalitypf:gradient}
    \| u \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} 
    \| \bff \|_{L^{p}(\Omega)}
    .
\end{align}
We call this inequality \emph{Poincar\'e--Friedrichs inequality} and the constant $C_{\grad,\Omega,p}$ is called the \emph{Poincar\'e--Friedrichs constant} with exponent $p$. 
The different scalar potentials only differ by a constant. 
Therefore, %when the domain is connected, then 
this inequality is equivalent to the following statement: 
\begin{align}\label{math:intro:poincarefriedrichs:grad}
    \min\limits_{c\in\bbR}
    \| u - c \|_{L^{p}(\Omega)}
    \leq 
    C_{\grad,\Omega,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    \forall
    u \in W^{1,p}(\Omega) % TODO: Use \forall notation 
    .\tag{PF} 
\end{align}
% Given a domain $\Omega \subseteq \bbR^{n}$ and Lebesgue exponent $1 \leq p \leq \infty$,
% we search for a constant $C_{\grad,\Omega,p} > 0$ such that 
% \begin{align}\label{math:intro:poincarefriedrichs:grad}
%     \min\limits_{c\in\bbR}
%     \| u - c \|_{L^{p}(\Omega)}
%     \leq 
%     C_{\grad,\Omega,p} \| \nabla u \|_{L^p(\Omega)},
%     \quad 
%     u \in W^{1,p}(\Omega) % TODO: Use \forall notation 
%     .\tag{PF} 
% \end{align}
Finding a potential of a given gradient $\nabla u$ is generally a nonlinear operation. 
The operator norm of any \emph{linear} potential construction is an upper bound for the Poincar\'e--Friedrichs constant. 
% Even though the existence of such a bounded linear potential construction is not evident a priori, picking the average-free potential 
One example for a bounded linear operator is the average-free potential:
\begin{align*}
    \Phi_{\grad}( \bff ) 
    := 
    \argmin\limits_{ \substack{ v \in W^{1,p}(\Omega) \\ \int_\Omega v = 0 } } \| \nabla v - \bff \|_{L^{p}(\Omega)},
    \quad 
    \forall 
    \bff \in \nabla W^{1,p}(\Omega)
    .
\end{align*} 
Its operator norm is the constant $C_{{\rm P},\Omega,p} > 0$ in the Poincar\'e inequality 
\begin{align}\label{math:intro:poincare:grad}
    \| u - u_{\Omega} \|_{L^{p}(\Omega)}
    \leq 
    C_{{\rm P},\Omega,p} \| \nabla u \|_{L^p(\Omega)},
    \quad 
    \forall 
    u \in W^{1,p}(\Omega)
    , \tag{P} % DONE: tag Poincar\'e inequality 
\end{align}
where $u_\Omega$ denotes the average of $u$.
Considerable research effort has gone into finding (or estimating) the optimal constants in the Poincar\'e--Friedrichs inequalities~\eqref{math:intro:poincarefriedrichs:grad} or Poincar\'e inequalities~\eqref{math:intro:poincare:grad}.

@@

% DONE: Introduce a reference simplex 
% DONE: DEFINE THE JACOBIAN 
\todo{MWL: If we need to define the Jacobian formally, where should that be put? Maybe an extra paragraph \textbf{Notation} immediately at the end of the introduction section.}

@@

\begin{lemma}    
    Suppose that an $n$-dimensional manifold triangulation $\calT$ has a shelling $T_{0}, T_{1}, T_{2}, \dots, T_{M}$.
    For $0 \leq m < M$, write 
    $$
        X_{m} := T_{0} \cup T_{1} \cup \dots \cup T_{m},
        \qquad 
        I_{m} := X_{m} \cap T_{m+1}.
    $$ 
    Then $I$ is the union of $k$ different facets of $T$, 
    and their intersection is a simplex $S \in \calT$ of dimension $n-k$.
    Moreover, 
    $$
        \patch_{X_{m+1}}(S) = \patch_{\calT}(S).
    $$
\end{lemma}
\begin{proof}
    We know $I_{m}$ is a submanifold of the boundary of $T$, and so it must be a collection of $k$ facets
    that constitute the local patch around some simplex $S$ of dimension $n-k$.
    To show that $\patch_{X_{m+1}}(S) = \patch_{\calT}(S)$, let us assume that $T \in \patch_{\calT}(S) \setminus \patch_{X_{m+1}}(S)$.
    By assumption, $S$ must be a boundary simplex of $X_{m+1}$. 
    \todo{Complete this here. The barycenter is contained in the interior of a boundary disk.}
\end{proof}

\begin{theorem}\label{theorem:starreflection}
    Let $\calT$ be an $n$-dimensional triangulation and let $S \in \calT$ be a simplex of dimension $k < n$.
    Let $T \in \patch_{\calT}(S)$ be of dimension $n$ and let $M := |\patch_{\calT}(S)| \setminus T$. 
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \varphi : T \rightarrow M
    \end{align*}
    satisfying
    \begin{align*}
        \Lip( \varphi,      T ) \leq \Cvier{n}{k} \diam(T),
        \quad 
        \Lip( \varphi^{-1}, M ) \leq \Cfive{n}{k} \diam(T),
    \end{align*}
    where $\Cvier{n}{k} \leq ...$ and $\Cfive{n}{k} \leq ...$
\end{theorem}
\begin{proof}
\begin{itemize}
    \item 
    \todo{Choose hyperplane that puts T on one side}
    We use Lemma~\ref{lemma:oppositesubsimplex}
    \item 
    \todo{Align barycenter of opposite subsimplex with coordinate axis}
    \item 
    \todo{opposite simplex cast to disk}
    \item 
    \todo{roll out rest of sphere onto disk}
    \item 
    \todo{estimate reflection}
\end{itemize}
\end{proof}

% https://mathoverflow.net/questions/232302/lipschitz-constant-of-cental-projection-of-unit-ball-to-surface-of-convex-body
% https://eudml.org/doc/258483
% https://math.stackexchange.com/questions/1476085/uniformly-convex-approximation-of-convex-domain
% SMOOTH EXHAUSTION FUNCTIONS IN CONVEX DOMAINS
\begin{lemma}
    Let $K \subseteq \bbR^{n}$ be a compact set star-shaped with respect to the origin 
    that contains the ball $B_{r}(0)$ and that is contained in the ball $B_{R}(0)$. Then the function 
    \begin{align*}
        f : S \rightarrow \partial K, \quad x \mapsto \arg( \partial K \cap \bbR x )
    \end{align*}
    satisfies 
    \begin{align*}
        R^{2}/r % euclidean 
    \end{align*}
    If $K$ is convex, then 
    \begin{align*}
        g : S \rightarrow \bbR, \quad x \mapsto \| \bbR x \cap \partial K \|
    \end{align*}
    satisfies 
    \begin{align*}
        R \sqrt{ R^{2} / r^{2} - 1 } % angular distance 
        \\
        \frac R r \sqrt{ R^{2} - r^{2} } % angular distance 
    \end{align*}
\end{lemma}
\begin{lemma}
    .
\end{lemma}

@@

@@

\section{Numerical examples}

\subsection{Two-dimensional examples}

\subsection{Three-dimensional examples}
